---
title: "Oneway ANOVA"
---

```{r}
#| include: false
library(tidyverse)

mystyle <- function(x){
    gt::gt(x) %>% 
    gt::opt_stylize(1, color = "blue") %>% 
    gt::tab_options(table.font.names = "Times New Roman") %>% 
    gt::tab_style(
        style = gt::cell_text(weight = "bold"),
        locations = gt::cells_column_labels()
        )
    }
```

## Background

Oneway analysis of variance is used when there are more than two levels of a predictor variable and
one dependent variable.

## Hypothesis

H~0~ - There is no difference in the means for each group

Ha - At least one of the means is significantly different from the others

## Assumptions

1.  **Normality** – Each sample should be drawn from a normally distributed population. It is not
    required for large sample sizes. If normality is violated use **Kruskal-Wallis** test
2.  **Equal Variance** - The variance of all the groups must be similar. If variances are equal, use
    **ANOVA**. If variances are not equal, use the **Welch ANOVA**
3.  **Independence** - The observation in each group must be independent from each other. They must
    also come from a random process.
4.  **Outlier** - Data should be devoid of significant outliers

## Data

We look at data from a nursery of newborns, comparing their weights after grouping their 5-minute
APGAR scores into Low (0-3), Medium(4-7), and High (8-10).

```{r}
df_babies <- 
    readxl::read_xlsx("C:/Dataset/mbu2.xlsx") %>% 
    filter(age == 0 & apgar2 <= 10 & sex != "Missing" & wt < 7.5) %>% 
    select(apgar5 = apgar2, bwt = wt, sex) %>% 
    drop_na() %>% 
    mutate(
        sex = factor(sex),
        apgar5 = case_when(
            apgar5 < 4 ~ "0-3",
            apgar5 < 8 ~ "4-7",
            apgar5 <= 10 ~ "8-10") %>% 
            factor(levels = c("0-3", "4-7", "8-10"), ordered = TRUE))

df_babies %>% 
    summarytools::dfSummary(graph.col = F)
```

Data summary and visualization

```{r}
df_babies %>% 
    group_by(apgar5) %>% 
    summarise(across(
        .cols = c(bwt), 
        .fns = list(
            "Mean" = ~mean(.x), 
            "SD" = ~sd(.x),
            "Variance" = ~var(.x),
            "N" = ~n()))) %>% 
    kableExtra::kable()
```

```{r}
#| label: fig-one
#| fig-cap: "Distriubution of birth weight for various APGAR categories"
#| fig-height: 3
#| fig-width: 5
df_babies %>% 
    ggplot(aes(y = bwt, x = apgar5)) + 
    geom_boxplot(alpha = .3) +
    labs(
        y = "Birth Weight (kgs)", 
        x = "APGAR",
        fill = "APGAR Category") +
    theme_light()
```

```{r}
#| label: fig-two
#| fig-cap: "Distriubution of birth weight for various APGAR categories"
#| fig-height: 3
#| fig-width:  6
df_babies %>% 
    ggplot(aes(y = bwt, x = apgar5)) + 
    geom_boxplot(alpha = .3) +
    labs(
        y = "Birth Weight (kgs)", 
        x = "APGAR") +
    theme_light()+
    theme(
        strip.text = element_text(face = "bold", size = 12),
        strip.background = element_rect(fill = "black"))+
    facet_wrap(. ~ sex)
```

## Fit model

All data together

```{r}
df_babies %>% 
    rstatix::anova_test(bwt ~ apgar5) %>% 
    kableExtra::kable()
```

Data grouped by sex

```{r}
df_babies %>% 
    group_by(sex) %>% 
    rstatix::anova_test(bwt ~ apgar5) %>% 
    kableExtra::kable()
```

## Mutltiple comparison

All data together

```{r}
df_babies %>% 
    rstatix::tukey_hsd(bwt ~ apgar5) %>% 
    kableExtra::kable()
```

Data grouped by sex

```{r}
df_babies %>% 
    group_by(sex) %>% 
    rstatix::tukey_hsd(bwt ~ apgar5) %>%  
    kableExtra::kable()
```

## Checking assumptions

### Plotting

This is a large sample size but it will be performed for practice

```{r}
#| fig-height: 10
#| fig-width: 8
df_babies %>% 
    aov(bwt ~ apgar5, data = .) %>% 
    performance::check_model()
```

There appear to be significant violations of the assumptions of the model

### Testing

#### Normality

```{r}
df_babies %>% 
    aov(bwt ~ apgar5, data = .) %>% 
    broom::augment() %>% 
    rstatix::shapiro_test(.resid) %>%
    mystyle()
```

#### Equality of variance

```{r}
df_babies %>% 
    rstatix::levene_test(bwt ~ apgar5)%>%
    mystyle()
```
