---
title: "Bland and Altman Analysis"
---

## Introduction

Correlation quantifies the degree to which two variables are related. However, a high correlation
does not automatically imply good agreement between the two methods. The correlation coefficient and
regression technique are sometimes inadequate and can be misleading when assessing agreement because
they evaluate only the linear association of two sets of observations. The r measures the strength
of a relation between two variables, not their agreement. Similarly, r2, named the coefficient of
determination, only tells us the proportion of variance the two variables have in common. Finally,
the significance test may show that the two methods are related, but it is evident that two methods
designed to measure the same variable are related.

## Case scenario

A Bland-Altman analysis, also known as a Bland-Altman plot or Tukey mean-difference plot compares
two different measurement techniques or instruments. It's commonly used in analytical chemistry and
biomedicine to assess the agreement between two assays or measurement methods.

Key Components: X-Axis: Displays the average of the measurements from the two methods.

Y-Axis: Displays the difference between the measurements from the two methods.

Mean Difference (Bias): A horizontal line representing the average difference between the two
methods.

Limits of Agreement: Lines representing the upper and lower limits of the differences, typically
calculated as the mean difference ± 1.96 times the standard deviation of the differences.

Interpretation: Bias: Indicates if there is a systematic difference between the two methods.

Limits of Agreement: Provides an estimate of the range in which 95% of the differences between the
two methods fall.

This analysis helps identify any systematic bias between the methods and assesses the typical range
of agreement. It's beneficial for determining if two measurement techniques can be used
interchangeably.

```{r}
#| include: false
library(tidyverse)
```

```{r}
df_poap <- 
    readxl::read_xlsx("C:/Dataset/pawper.xlsx") %>% 
    mutate(
        weight = parse_number(weight), 
        id = row_number()) %>% 
    filter(id != 343) %>% 
    select(-id)

```

Exploration

```{r}
#| fig-height: 4
#| fig-width: 7
df_poap %>% 
    DataExplorer::plot_histogram(
        ggtheme = theme_bw(),
        geom_histogram_args = list(
            fill = "steelblue", 
            color = "blue", 
            bins = 12)) 
```

```{r}
#| fig-height: 4
#| fig-width: 6

df_poap %>% 
    ggplot(aes(x = weight, y = pxl_wgt))+
    geom_point()+
    labs(x = "Actual weight (kgs)", y = "Pawper-XL weight (kgs)")+
    theme_bw()
```

```{r}
#| fig-height: 4
#| fig-width: 6

df_temp <- 
    df_poap %>% 
    select(pxl_wgt, weight) %>% 
    drop_na() %>% 
    mutate(
        means = (pxl_wgt + weight)/2,
        diffs = (pxl_wgt - weight),
        bias = mean(diffs),
        sd_diffs = sd(diffs),
        upper_loa = bias + 1.96 * sd_diffs,
        lower_loa = bias - 1.96 * sd_diffs) 
ula = mean(df_temp$upper_loa)
lla = mean(df_temp$lower_loa)
mn = mean(df_temp$bias)
xx = max(df_temp$means)

df_temp %>% 
    ggplot(aes(x = means, y = diffs)) +
    geom_point()+
    geom_hline(
        aes(yintercept = upper_loa),
        linetype = 2, color = "#2369BD")+
     geom_hline(aes(yintercept = lower_loa),
        linetype = 2, color = "#2369BD") +
    geom_hline(aes(yintercept = bias),
        linetype = 2, color = "#2369BD") +
    annotate(
        geom = 'text', 
        label = 'mean+1.96*SD', 
        x = xx, y = ula, hjust = 1, vjust = -1, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = str_glue({round(ula, 2)}), 
        x = xx, y = ula, hjust = 1, vjust = 1.2, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = 'mean-1.96*SD', 
        x = xx, y = lla, hjust = 1, vjust = -1, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = str_glue({round(lla, 2)}), 
        x = xx, y = lla, hjust = 1, vjust = 1.2, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = 'Bias', 
        x = xx, y = mn, hjust = 1, vjust = -1, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = str_glue({round(mn, 2)}), 
        x = xx, y = mn, hjust = 1, vjust = 1.2, 
        color = "#2369BD", size = 4, family = "serif")+
    labs(
        x = "Mean of Pawper-XL and actual weight (kgs)",
        y = "Difference")+
    theme_classic()
```

```{r}
#| fig-height: 4
#| fig-width: 6


df_temp <- 
    df_poap %>% 
    select(pxl_wgt, weight) %>% 
    drop_na() %>% 
    mutate(
        means = (pxl_wgt + weight)/2,
        diffs = (pxl_wgt - weight)/weight*100,
        bias = mean(diffs),
        sd_diffs = sd(diffs),
        upper_loa = bias + 1.96 * sd_diffs,
        lower_loa = bias - 1.96 * sd_diffs) 
ula = mean(df_temp$upper_loa)
lla = mean(df_temp$lower_loa)
mn = mean(df_temp$bias)
xx = max(df_temp$means)

df_temp %>% 
    ggplot(aes(x = means, y = diffs)) +
    geom_point()+
    geom_hline(
        aes(yintercept = upper_loa),
        linetype = 2, color = "#2369BD")+
     geom_hline(aes(yintercept = lower_loa),
        linetype = 2, color = "#2369BD") +
    geom_hline(aes(yintercept = bias),
        linetype = 2, color = "#2369BD") +
    annotate(
        geom = 'text', 
        label = 'mean+1.96*SD', 
        x = xx, y = ula, hjust = 1, vjust = -1, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = str_glue({round(ula, 2)}), 
        x = xx, y = ula, hjust = 1, vjust = 1.2, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = 'mean-1.96*SD', 
        x = xx, y = lla, hjust = 1, vjust = -1, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = str_glue({round(lla, 2)}), 
        x = xx, y = lla, hjust = 1, vjust = 1.2, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = 'Bias', 
        x = xx, y = mn, hjust = 1, vjust = -1, 
        color = "#2369BD", size = 4, family = "serif")+
    annotate(
        geom = 'text', 
        label = str_glue({round(mn, 2)}), 
        x = xx, y = mn, hjust = 1, vjust = 1.2, 
        color = "#2369BD", size = 4, family = "serif")+
    labs(
        x = "Mean of Pawper-XL and actual weight (kgs)",
        y = "Percentage Difference")+
    theme_classic()
```

```{r}
library(magrittr)
df_poap %$% 
    BlandAltmanLeh::bland.altman.stats(
        group1 = weight, group2 = pxl_wgt)
```
