---
title: "Exploratory Data Analysis"
---

This chapter introduces the initial graphical exploration of data using the `DataExplorer` package.
We begin by importing the babies data.

```{r}
#| message: false
#| include: false

library(tidyverse)
```

Note that the following are very useful functions that will not be executed in this page

`DataExplorer::create_report()`

`summarytools::dfSummary() %>% summarytools::stview()`

`SmartEDA::ExpReport()`

We begin by reading in the data and retaining

```{r}
df_schisto <- 
    readxl::read_xlsx("C:/Dataset/Schisto.xlsx") %>%
    janitor::clean_names() %>%
    rename(
        weight = q35weight,
        height = q36height,
        sex = q4sex,
        religion = q6religion, 
        schisto = sh,
        educ_status = educationalstatus,
        ageyrs = q3age) %>% 
    mutate(
        across(where(is.character), as.factor), 
        schisto = case_when(
            schisto == "+" ~ "+", schisto == "++" ~ "++",
            schisto == "+++" ~ "+++", schisto == "++++" ~ "++++", 
            TRUE ~ "nil") %>% 
            factor(levels = c("nil", "+", "++", "+++", "++++"))) %>% 
    select(
        serial_no, ageyrs, sex, religion, educ_status, 
        height, weight, hb, wbc, schisto) 
```

## General overview

We begin with a general overview of the data

```{r}
df_schisto %>% glimpse()
```

And then

```{r}
df_schisto %>% summary()
```

And the use the psych package

```{r}
df_schisto %>% 
    psych::describe() %>% 
    gt::gt()
```

And then plot this overview

```{r}
df_schisto %>% 
    DataExplorer::plot_intro()
```

```{r}
df_schisto %>% 
    summarytools::dfSummary()
```

```{r}
df_schisto %>% 
    dlookr::describe() %>% 
    flextable::flextable()
```

And using the `psych` package

```{r}
df_schisto %>% 
    psych::describe()%>% 
    flextable::flextable()
```

## Missing data

Next, we derive explore the missing data. The plot outlines the percentage of missing data for each
with the legend showing if the number of missing is good, ok or bad.

```{r}
df_schisto %>% 
    DataExplorer::plot_missing()
```

Also,

```{r}
df_schisto %>% inspectdf::inspect_na()
```

```{r}
df_schisto %>% 
    inspectdf::inspect_na() %>% 
    inspectdf::show_plot()
```

## Categorical variables

We begin by exploring the categorical variables

```{r}
df_schisto %>% 
    dlookr::diagnose_category() %>% 
    flextable::flextable()
```

Next we visualise categorical variable with a barplot

```{r}
df_schisto %>% 
    DataExplorer::plot_bar()
```

## Continuous variables

Now to continuous variables

```{r}
df_schisto %>% 
    dlookr::diagnose_numeric() %>% 
    flextable::flextable()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_histogram()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_boxplot(by = "sex")
```

```{r}
df_schisto %>% 
    DataExplorer::plot_density()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_qq()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_qq( by = "sex", )
```

## Outliers

```{r}
df_schisto %>% 
    dlookr::diagnose_outlier() %>% 
    flextable::flextable()
```

## Correlation

Next, we look out for correlation in the continuous variables

```{r}
df_schisto %>% 
    drop_na() %>% 
    DataExplorer::plot_correlation(type = "continuous")
```

```{r}
df_schisto %>% 
    inspectdf::inspect_cor() %>% 
    gt::gt()
```

```{r}
df_schisto %>% 
    inspectdf::inspect_cor() %>%
    inspectdf::show_plot()
```

```{r}
numeric_df <- 
    df_schisto %>%
    select(where(is.numeric)) 

PerformanceAnalytics::chart.Correlation(
    numeric_df,
    histogram = T,
    pch = 12
)
```

## Scatterplots

Next is scatterplots

```{r}
df_schisto %>% 
    DataExplorer::plot_scatterplot(by = "weight")
```
