---
title: "Exploratory Data Analysis"
---

This chapter introduces the initial graphical exploration of data using the `DataExplorer` package.
We begin by importing the babies data.

```{r}
#| message: false
#| include: false


library(tidyverse)
```

```{r}
df_schisto <- 
    readxl::read_xlsx("C:/Dataset/Schisto.xlsx") %>%
    janitor::clean_names() %>% 
    mutate(
        across(where(is.character), as.factor),
        mcv = as.numeric(mcv)) %>% 
    select(-c(q5tribe, q1date, serial_no))
```

We begin with a general overview of the data

```{r}
df_schisto %>% 
    DataExplorer::introduce() %>% 
    t()
```

And then plot this overview

```{r}
df_schisto %>% 
    DataExplorer::plot_intro()
```

```{r}
df_schisto %>% 
    dlookr::describe() %>% 
    flextable::flextable()
```

## Missing data

Next, we derive explore the missing data. The plot outlines the percentage of missing data for each
with the legend showing if the number of missing is good, ok or bad.

```{r}
df_schisto %>% 
    DataExplorer::plot_missing()
```

```{r}
# df_schisto %>% dlookr::plot_na_hclust()
# df_schisto %>% dlookr::plot_na_pareto()
# df_schisto %>% dlookr::plot_na_intersect()
```

## Categorical variables

We begin by exploring the categorical variables

```{r}
df_schisto %>% 
    dlookr::diagnose_category() %>% 
    flextable::flextable()
```

Next we visualise categorical variable with a barplot

```{r}
df_schisto %>% 
    DataExplorer::plot_bar()
```

## Continuous variables

Now to continuous variables

```{r}
df_schisto %>% 
    dlookr::diagnose_numeric() %>% 
    flextable::flextable()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_histogram()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_boxplot(by = "q4sex")
```

```{r}
df_schisto %>% 
    DataExplorer::plot_density()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_qq()
```

```{r}
df_schisto %>% 
    DataExplorer::plot_qq( by = "q4sex", )
```

## Outliers

```{r}
df_schisto %>% 
    dlookr::diagnose_outlier() %>% 
    flextable::flextable()
```

```{r}
# df_schisto %>% dlookr::plot_outlier()
```

## Correlation

Next, we look out for correlation in the continuous variables

```{r}
df_schisto %>% 
    drop_na() %>% 
    DataExplorer::plot_correlation(type = "continuous")
```

## Scatterplots

Next is scatterplots

```{r}
df_schisto %>% 
    DataExplorer::plot_scatterplot(by = "q35weight")
```
