---
title: "Barplot"
---

```{r}
#| include: false

rm(list=ls(all.names=TRUE))
library(tidyverse)
```

To start we create a fictitious data

```{r comment = NA}
df <- 
  data.frame(
      dose = c("D0.5", "D1", "D2"), 
      len = c(4.2, 10, 29.5))
df
```

## Basic barplot

Next, we make the basic Barplot

```{r}
p <- 
  df %>% 
  ggplot(aes(x = dose, y = len)) 

p + geom_bar(stat="identity")
```

Next we flip the barplot horizontal, change the size of the bar width and change the theme

```{r comment = NA}
p + 
  geom_bar(
      stat = "identity", 
      width = 0.8, 
      color = "blue", 
      fill = "grey90") +
  coord_flip() +
  theme_bw()
```

Next we limit the observations to just two

```{r comment = NA}
p + 
  geom_bar(
      stat = "identity", 
      width = 0.8, 
      color = "black", 
      fill = "steelblue") +
  scale_x_discrete(limits=c("D0.5", "D2"))
```

Next we put labels on the bars at the outside and inside

```{r comment = NA}
p +
    geom_bar(stat="identity", fill="steelblue")+
    geom_text(
        aes(label=len), 
        vjust=-0.5, 
        size=4, 
        col = "black")+
    theme_minimal()
```

```{r comment = NA}
# Change barplot line colors by groups
p <- 
    ggplot(df, aes(x = dose, y = len, color = dose)) + 
    geom_bar(stat = "identity", fill = "white") + 
    geom_text(aes(label=len), vjust=1.5, size=5, col = "red")
p
```

## Use custom color palettes

```{r comment = TRUE}
p + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```

Use brewer color palettes

```{r comment = TRUE}
p + scale_color_brewer(palette="Dark2")
```

```{r comment = TRUE}
p + scale_color_grey() + theme_classic()
```

Change barplot fill colors by groups

```{r comment = TRUE}
# 
p <- 
  ggplot(df, aes(x=dose, y=len, fill=dose)) +
  geom_bar(stat="identity")+theme_minimal()
p
```

Use custom color palettes

```{r comment = NA}
p + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```

```{r comment = NA}
p + 
  scale_fill_brewer(palette="Dark2")
```

Use grey scale

```{r comment = TRUE}
p + scale_fill_grey()
```

```{r comment = NA}
ggplot(df, aes(x=dose, y=len, fill=dose))+
	geom_bar(stat="identity", color="black")+
	scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  theme_minimal()
```

Change bar fill colors to blues

```{r comment = TRUE}
p + scale_fill_brewer(palette="Blues")
	p + theme(legend.position="top")
	p + theme(legend.position="bottom")
	p + theme(legend.position="none")
```

Change position of bars

```{r comment = TRUE}
p + 
  scale_x_discrete(limits=c("D2", "D0.5", "D1"))
```

```{r comment = NA}
dat <- foreign::read.dta("C:/dataset/bea_organ_damage_28122013.dta")
BC <- 
  dat %>% 
  select(q2idtype, q3sex) %>% 
  na.omit() %>% 
  group_by(q2idtype, q3sex) %>% 
  summarize(Freq = n()) %>% 
  ggplot(aes(x=q2idtype, y=Freq, fill=q3sex))
```

Next we draw the barplot using the economist theme from the `ggthemes` package

```{r comment = NA}
BC +  
  geom_bar(stat="identity", position= position_dodge()) +
	geom_text(aes(label=Freq), vjust=1.6, color="black", 
	          size=4, position = position_dodge(0.9)) +
	scale_fill_brewer(palette="Reds") + 
	labs(title="My Barplot", x="Case or Control", y="Frequency") +
	scale_color_discrete(name="Sex") + 
	ggthemes::theme_stata()
```

Next we plot a baroplot with error bars. To do that we first we form the `ggplot` object that we
call *BC*.

```{r comment = NA}
BC <-
  dat %>% 
  select(Type = q2idtype, Sex = q3sex, q12weight) %>% 
  na.omit() %>% 
  group_by(Type, Sex) %>% 
  summarize(Mean.wgt = mean(q12weight), SD.wgt = sd(q12weight)) %>% 
  ggplot(aes(x=Type, y=Mean.wgt, fill=Sex))
```

And then plot the graph

```{r comment = NA}
BC +  
    geom_bar(stat="identity", position=position_dodge()) + 
    geom_errorbar(
        aes(ymin = Mean.wgt-SD.wgt, ymax = Mean.wgt+SD.wgt), 
	    width=.2, 
        size=0., 
        position=position_dodge(.9)) +
	labs(
	    title="Mean weight with error bars", 
	    x="Case or Control", 
	    y="Mean(kgs)") +
	scale_fill_brewer(palette="Paired") + 
	ggthemes::theme_stata()
```

```{r comment = TRUE}

df2 <- data.frame(supp=rep(c("VC", "OJ"), each=3),
                dose=rep(c("D0.5", "D1", "D2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))
head(df2)
```

```{r}
df2 %>% 
  ggplot(aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity")
```

```{r}
df2 %>% 
  ggplot(aes(x=dose, y=len, fill=supp)) + 
  geom_bar(stat="identity", position=position_dodge())
```

Change color manually

```{r comment = NA}
p <- 
  ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  scale_fill_manual(values=c('#999999','#E69F00')) +
  theme_minimal()
p
```

Create some data

```{r comment = NA}
df_sorted <- 
  tibble(supp = factor(rep(c("VC", "OJ"), each=3)),
         dose = rep(c("0.5", "1", "2"),2),
         len = c(6.8, 15, 33, 4.2, 10, 29.5)) %>% 
  arrange(dose, supp) %>% 
  group_by(dose) %>% 
  mutate(label_ypos=cumsum(len))
df_sorted
```

```{r comment = NA}
df_sorted %>% 
  ggplot(aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity")+
  geom_text(aes(y=label_ypos, label=len), vjust=1.6, 
            color="white", size=3.5)+
  scale_fill_brewer(palette="Paired")+
  theme_minimal()
```

Plotting barplot with x-axis treated as continuous variable

```{r comment = NA}
# 
df_sorted %>% 
  mutate(dose = as.numeric(dose)) %>% 
  ggplot(aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity", position=position_dodge())+
  scale_fill_brewer(palette="Paired")+
  theme_minimal()
```

```{r comment = NA}
# Axis treated as discrete variable
df_sorted %>% 
  mutate(dose = as.factor(dose)) %>% 
  ggplot(aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity", position=position_dodge())+
  scale_fill_brewer(palette="Paired")+
  theme_minimal()
```

```{r comment = NA}
ToothGrowth %>% 
  mutate(dose = as.factor(dose)) %>% 
  group_by(supp, dose) %>% 
  summarise(sd = sd(len), len = mean(len), .groups = "drop") %>% 
  ggplot(aes(x=dose, y=len, fill=supp)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2, 
                position=position_dodge(.9)) +
  labs(title="Plot of length  per dose", x="Dose (mg)", y = "Length")+
  scale_fill_brewer(palette="Paired") + 
  theme_minimal()
```

## Plotting raw data

```{r}
babies <- 
    dget("babies_clean") %>% 
    select(apgar5cat, died)

babies %>% 
    ggplot(aes(x = apgar5cat, fill = apgar5cat)) +
    geom_bar(stat = 'count') +
    scale_fill_grey(
        name = "APGAR Category",
        label = c("LOW", "MEDIUM", "HIGH"))+
    theme_bw()
```
