---
title: "Single Continuous Variable"
---

```{r}
#| include: false
#| message: false

library(tidyverse)
```

This section will demonstrate the use of simple linear regression to describe the relationship
between variables. We begin by importing the `blood` data and summarizing it as below:

```{r}
#| message: false
df_blood <- 
    read_csv("C:/Dataset/blood.csv") %>% 
    select(hb, hct)

df_blood %>% 
    summarytools::dfSummary(graph.col = F, labels.col = F)
```

## Plotting

We begin by plotting the distribution of the variables involved

```{r}
#| label: fig-four
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "Relationship between hemoglobin and hematocrit"


df_blood %>% 
    pivot_longer(cols = c(hb, hct)) %>% 
    ggplot(aes( x = value)) +
    geom_histogram(bins = 8, fill = "gold", color = "black")+
    facet_wrap(facets = "name", scales = "free") +
    theme_bw()
```

Next, we plot the relationship between the `hct`and `hb` variables and note the linear relationship.

```{r}
#| label: fig-one
#| fig-height: 4
#| fig-width: 5
#| fig-cap: "Relationship between hemoglobin and hematocrit"


df_blood %>% 
    ggplot(aes(x = hct, y = hb)) +
    geom_point()+
    geom_smooth(formula = y ~ x, method = "lm", se = FALSE)+
    theme_bw()
```

## Assumptions

1.  **Linearity**: The relationship between the independent variable (X) and the dependent
    variable (Y) is linear.
2.  **Independence**: The observations are independent of each other.
3.  **Homoscedasticity**: The residuals (errors) have constant variance at every level of X.
4.  **Normality**: The residuals of the model are normally distributed.
5.  **No multicollinearity**: This is usually more relevant for multiple regression, but it means
    the independent variables aren't too highly correlated with each other.

## Model fitting

```{r}
model <- 
    df_blood %>% 
    lm(hb ~ hct, data = .)
```

## Visualising model

### R base `summary`

```{r}
model %>% summary()

```

### `tab_model`

```{r}
model %>% sjPlot::tab_model() 
```

### `tidy`

```{r}
model %>% broom::tidy() %>% kableExtra::kable()
```

### `tbl_uvregression`

```{r}
df_blood %>% 
    gtsummary::tbl_uvregression(
        y = hb,
        method = "lm"
    )
```

## Checking Assumptions

We see no significant violation of the model assumptions

```{r}
#| label: fig-two
#| fig-height: 8
#| fig-width: 7
#| fig-cap: "Model Assumptions of simple linear regression"


performance::check_model(model)
```

## Prediction interval

```{r}
#| label: fig-three
#| fig-cap: Relationship between HB4 and HCT4 with fillted line, prediction and se intervals"
#| fig-width: 6
#| fig-height: 5
#| warning: false


model %>% 
    predict(interval = "predict") %>% 
    as_tibble() %>% 
    bind_cols(df_blood) %>% 
    ggplot(aes(x = hct, y = hb)) +
    geom_point() +
    geom_smooth(method = "lm", formula = y~x, se=T)+
    geom_line(aes(y = lwr), col = "coral2", linetype = "dashed") +
    geom_line(aes(y = upr), col = "coral2", linetype = "dashed") +
    labs(
        x = "HCT (%)", 
        y = "HB (mg/dl)", 
        caption = "Nurse Data 2015")+
    theme_bw()
```

```{r}
model %>% 
    emmeans::emmeans(~hct, at = list(hct = c(20, 25, 30, 35)))
```

## Report

```{r}
report::report(model)
```
