---
title: "Single Categorical Predictor"
---

This segment deals with the basic logistic regression using a single categorical predictor. We use
the `babies`data. This is summarised first-MinuteFifth-minute below:

```{r}
#| message: false
#| echo: false

library(tidyverse)
library(magrittr)

babies <- 
    readxl::read_xlsx("C:/Dataset/babies.xlsx") %>% 
    rename(apgar5 = apgar2) %>% 
    mutate(
        across(c(sex, hosp, delmode), ~factor(.x)),
        died = case_when(
            outcome == "Died" ~ "Yes",
            outcome == "Discharged" ~ "No") %>%  factor(),
        apgar1cat = case_when(
            apgar1 < 4 ~ "Low",
            apgar1 < 7 ~ "Medium",
            apgar1 >= 7 ~ "High") %>% 
            factor(levels = c("Low", "Medium", "High")),
         apgar5cat = case_when(
            apgar5 < 4 ~ "Low",
            apgar5 < 7 ~ "Medium",
            apgar5 >= 7 ~ "High")%>% 
            factor(levels = c("Low", "Medium", "High"))) %>% 
        select(-c(outcome, id)) %>% 
    labelled::set_variable_labels(
        hosp = "Hospital",
        sex = "Sex of baby",
        wt = "Weight of baby",
        apgar1 = "First Minute APGAR Score",
        apgar5 = "Fifth Minute APGAR Score", 
        delmode = "Mode of Delivery",
        died = "Mortality",
        apgar1cat = "First minute APGAR Category",
        apgar5cat = "Fifth minute APGAR Category") 


babies <- 
    dget("babies_clean") %>% 
    select(apgar5cat, died)

gtsummary::theme_gtsummary_eda()
gtsummary::theme_gtsummary_compact()

babies %>% 
    gtsummary::tbl_summary() %>% 
    gtsummary::bold_labels()
```

## Research question

-   What is the association between death and the fifth-minute APGAR Score categories?
-   What are the predicted probabilities of death for the various fifth minute APGAR score
    categories?
-   Do these probabilities differ significantly from each other?

## Graphing variables

```{r}
#| label: fig-one
#| fig-cap: "Barplot of Five minute APGAR categories and Mortality"
#| fig-width: 6
#| fig-height: 4
babies %>% 
    group_by(apgar5cat, died) %>% 
    summarize(count = n(), .groups = "drop") %>% 
    group_by(apgar5cat) %>% 
    mutate(perc = count/sum(count)) %>% 
    ggplot(
        aes(
            x = apgar5cat, 
            y = count, 
            fill = died, 
            label = paste0(count, "\n (", scales::percent(perc), ")"))) +
    geom_bar(stat = "identity", position = position_dodge())+
    labs(x = "Five minute APGAR", y = "Frequency", fill = "Mortality")+
    geom_text(
        vjust = -.25, 
        color= "black", 
        size = 3, 
        fontface="italic", 
        position = position_dodge(width = 1))+
    ylim(c(0, 6000)) +
    scale_fill_manual(values = c("#70161E","#A4B6AC"))+
    theme_bw()
    
```

Alternatively, we can use the fifth-minute

```{r}
#| label: fig-two
#| fig-cap: "Distribution of Five minute APGAR categories and Mortality"
#| fig-width: 6
#| fig-height: 4
babies %$% 
    sjPlot::plot_xtab(
        grp = died, x = apgar5cat, margin = "row", 
        bar.pos = "stack", show.summary = TRUE, 
        coord.flip = TRUE)+
    theme_bw()
    
```

## Regression

```{r}
model <- 
    babies %>% 
    glm(died ~ apgar5cat, family = binomial, data = .)

summary(model)
```

```{r}
babies %>% 
    select(died, apgar5cat) %>% 
    gtsummary::tbl_uvregression(
        method = glm,
        y = died,
        method.args = family(binomial),
        exponentiate = TRUE,
        pvalue_fun = function(x) gtsummary::style_pvalue(x, digits = 3)) %>% 
    gtsummary::modify_header(
            estimate ~ "**OR**",
            label ~ "**Variable**") %>%
    gtsummary::bold_labels()
```

## Model evaluation

```{r}
model %>% performance::check_model(residual_type = "normal")
```

## Predicted probabilities

```{r}
model %>% 
    ggeffects::ggeffect() %>% 
    as_tibble() %>% 
    unnest(cols = c(apgar5cat)) %>% 
    kableExtra::kable()
```

## Plotting effects & predicted probabilities

```{r}
#| fig-width: 6
#| fig-height: 4
#| fig-align: center

model %>% 
    sjPlot::plot_model(type = "pred", terms = "apgar5cat") +
    theme_bw()
```

```{r}
#| fig-width: 6
#| fig-height: 4
#| fig-align: center

model %>% 
    sjPlot::plot_model(
        type = "est", 
        show.p = TRUE, 
        show.values = T, 
        show.intercept = T) +
    theme_bw()
```

## Pairwise comparison

```{r}
babies %>% 
    select(died, apgar5cat) %>% 
    gtsummary::tbl_uvregression(
        method = glm,
        y = died,
        method.args = family(binomial),
        exponentiate = TRUE,
        pvalue_fun = function(x) gtsummary::style_pvalue(x, digits = 3),
        add_pairwise_contrast = TRUE,
        contrast_adjust = "holm") %>% 
    gtsummary::modify_header(
            estimate ~ "**OR**",
            label ~ "**Variable**") %>%
    gtsummary::bold_labels()
```
