---
title: "`nlme` package"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls(all=TRUE))
library(tidyverse)
```

```{r}
cint <- dget("C:/Dataset/cint_data_11042021")
```

Since this is a large dataset we start off by selecting the variables we need, convert it to the
long format, and generate the "Time" variable

```{r comment=NA}
df <-
    cint %>% 
    select(
        sid, cca_0, cca_12, sex, ageyrs, resid, physical, 
        hpt_old, cart) %>% 
    gather(period, cca, cca_0:cca_12) %>% 
    mutate(
        Time = ifelse(
            period == "cca_0", 0, 
            ifelse(period == "cca_12", 1, NA)))
```

Here we begin to pick up abnormal or suspicious data escpecially in the cca. One way is to determine
the changes with time

```{r comment=NA}
df2 <- 
    df %>% 
    arrange(sid, period) %>%
    group_by(sid) %>% 
    mutate(
        cca_prev = lag(cca, order_by = sid),
        cca_diff = cca - cca_prev) 

df2 %>% 
    arrange(sid, period) %>% 
    head(10) %>% 
    kableExtra::kable()
```

And then plot the differences

```{r}
df2 %>% 
    na.omit() %>% 
    ggplot(aes(x = cca_diff)) +
    geom_boxplot()
```

Comment: Generally there was decrease in the cca values over the two time periods. Next we summarise
and visualise the data

```{r}
df %>% 
    na.omit() %>% 
    ggplot(aes(x = Time, y = cca, group = sid))+
    geom_jitter(show.legend = F, width = .01) +
    stat_smooth(
        formula = y ~x, 
        method = "lm", 
        se = FALSE, 
        col = "grey") +
    theme_light()
```

Next the UNCONDITIONAL NULL MODEL to evaluate clustering, thereby finding out if we need to even do
multilevel analysis

```{r comment=NA}
n0 <- 
    nlme::gls(
        cca ~ 1, data = df, na.action = na.omit, method = "ML")

n0 %>% 
    broom.mixed::tidy(conf.int=T) %>% 
    kableExtra::kable()
```

```{r comment=NA}
n1 <- 
    nlme::lme(
        cca ~ 1, random = ~ 1 | sid, 
        data = df, na.action = na.omit, method = "ML")

n0 %>% broom.mixed::tidy(conf.int=T) %>% kableExtra::kable()
```

```{r comment=NA}
n1 %>% reghelper::ICC()

anova(n0, n1)
```

Since ICC \> 0.05 and also the random effects CI does not include 0 there exist significant
clustering to suggest we do multilevel modeling. Next we test the UNCONDITIONAL SLOPE MODEL using
the FIXED slope

```{r comment=NA}
n2 <- 
    nlme::lme(
        cca ~ Time, random = ~ 1 | sid, 
        data = df, 
        na.action = na.omit, 
        method = "ML")

n2 %>% 
    broom.mixed::tidy(conf.int=T) %>% 
    kableExtra::kable()

n2 %>% reghelper::ICC()

anova(n0, n1, n2)
```

Next we test the UNCONDITIONAL SLOPE MODEL using the RANDOM slope

```{r comment=NA}
n3 <- 
    nlme::lme(
        cca ~ Time, 
        random = ~ Time | sid, 
        data = df, 
        na.action = na.omit, 
        method = "ML")


n3 %>% summary()

n3 %>% nlme::intervals(which = "fixed")

n3 %>% reghelper::ICC()

anova(n0, n1, n2, n3)
```

Next we test the FULL MODEL using the RANDOM slope

```{r comment=NA}
n4 <- 
    nlme::lme(
        cca ~ Time + hpt_old + sex + scale(ageyrs) + cart, 
        random = ~ 1 | sid, 
        data = df, 
        method = "ML", 
        na.action = na.omit)

n4 %>% 
    broom.mixed::tidy() %>% 
    kableExtra::kable()

n4 %>% reghelper::ICC()

anova(n0, n1, n2, n3, n4) 
```
