---
title: "`gtsummary`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtsummary)
```

# Reading in the data

```{r comment=NA}
df_cint_all <- dget("C:/Dataset/cint_data_clean")
df_mbu <- readxl::read_xlsx("C:/Dataset/mbu.xlsx")
```

# Formatting and cleaning the cint data

```{r comment=NA}
df_cint <- 
    df_cint_all %>%
    select(sex, bulb_0, bulb_12, bmi, sbp, dbp) %>% 
    mutate(
      bmicat = case_when(bmi < 25 ~ "Normal", bmi > 25 ~ "High") %>% 
        factor(levels = c("Normal", "High")),
      hpt = case_when(
        sbp > 120 | dbp > 80 ~ "Yes", sbp <= 120 & dbp <= 80 ~ "No" ) %>% 
        factor()) %>% 
    labelled::set_variable_labels(
        sex = "Sex",
        bulb_0 ="Bulb diameter at time 0",
        bulb_12 ="Bulb diameter at 12 months",
        bmicat = "Categorized BMI",
        sbp = "Systolic Blood Pressure",
        dbp = "Diastolic Blood Pressure",
        bmi = "Body Mass Index",
        hpt = "Hypertension present")
```

# Formatting and cleaning the MBU data

Yes it is here, and maybe here too

```{r}
df_mbu_clean <-
    df_mbu %>% 
    select(-c(pt_name, starts_with("diag"), address, Field1, ip, disweight)) %>% 
    mutate(wt = ifelse(wt > 8, NA, wt),
           age = ifelse(age > 60, NA, age),
           sex = ifelse(sex == "Missing", NA, sex) %>% factor(),
           place = ifelse((place == "Missing" | place == "Self"), NA, place),
           apgar1 = ifelse(apgar1 >10, NA, apgar1),
           apgar5 = ifelse(apgar2 >10, NA, apgar2), 
           dura_adm = difftime( disDateTime, admDateTime,units = "days"),
           dura_adm = ifelse((dura_adm <=0 | dura_adm > 90), NA, dura_adm),
           died = case_when(outcome == "Discharged" ~ "No",
                            outcome == "Died" ~ "Yes") %>% factor(),
           gestage = ifelse(gestage < 26 |gestage > 48, NA, gestage),
           del = case_when(del == "SVD" ~ "SVD",
                           del == "C/S" ~ "C/S",
                           del == "Forceps" ~ "Vacuum",
                           del == "Vacuum" ~ "Vacuum"))%>% 
    select(-c(apgar2, admDateTime, disDateTime, delDateTime, outcome)) %>% 
    drop_na() %>% 
    labelled::set_variable_labels(
      wt = "Weight (kgs)",
      sex = "Sex",
      age = "Age (days)",
      apgar1 = "APGAR (min 1)",
      apgar5 = "APGAR (min 5)",
      gestage = "Gestational Age",
      place = "Place of birth",
      del = "Mode of delivery",
      dura_adm = "Admission duration",
      died = "mortality")

df_mbu_clean %>% 
  summarytools::dfSummary(graph.col = F)
```

# Summarizing data

```{r comment=NA}
df_cint %>% 
  summarytools::dfSummary(graph.col = F, labels.col = F)
```

# Summary statistics tables

## Table 1

```{r comment=NA}
df_cint %>% 
    tbl_summary(
        by = sex,                    # aggregate table by sex
        missing_text = "(Missing)",  # Label missing data as such
        type = sbp ~ "continuous2",  # Report sbp with 2 or more statistics
        statistic = list(sbp ~ c("{mean},({sd})",
                                 "({min},{max})"),
                         bmicat ~ "{n}/{N} ({p}%)",
                         bulb_0 ~ c("{mean} ({sd})")),
        label = bmi ~ "BMI (Kg/m sq.)",   # To modify labels
        digits = dbp ~ 1) %>%         # Force dbp to have one decimal
    add_overall(last = T) %>%           # Add overall column
    modify_spanning_header(all_stat_cols() ~ "**Sex of Participants**") %>% 
    add_p(pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>%    # Adds p-value column
    add_q(method = "fdr",                   #Add p-value for multiple comparison
          pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
    add_stat_label() %>%        # Add specific stats to each variable
    add_n() %>%                     # Add valid observation to each variable
    bold_p() %>%       # Bold significant p-values
    add_significance_stars()   # add significance stars
```

## Table 2

```{r comment=NA}
df_mbu_clean %>% 
    tbl_summary(by = died) %>% 
    add_overall(last = T) %>%
    add_p() %>% 
    modify_spanning_header(all_stat_cols() ~ "**Mortality**") %>% 
    modify_caption(caption = "**Table 2**: MBU data by outcome") %>% 
    bold_labels()
```

## Table 3

This is table 3 of many😉

[^gt-summary-tables-1]

[^gt-summary-tables-1]: Need to insert this k3k3

```{r comment=NA}
df_cint %>% 
    tbl_summary(
        by = sex,                    # aggregate table by sex
        statistic = all_continuous() ~ "{mean} ({sd})") %>% # Stats touse for all continuous variables
    add_stat_label(label = all_continuous()~ "Mean(StD)") %>% # Label to give statistics
    add_difference() %>%    # Add a difeference, ci and p-value column
    modify_spanning_header(all_stat_cols()~ "**Sex**") %>%  # Add spanning header
    modify_caption("**Table 1. Patient Characteristics**") %>% # Add table title
    italicize_levels() %>%  # Italics for the levels
    bold_labels()   # Bold fo rhe labels

```

## Stratified table

```{r}
df_cint %>% 
    drop_na(bmicat) %>% 
    tbl_strata(
        strata = bmicat, ~.x %>%      # Add a strata to the table
            tbl_summary(
                by = sex
            ) %>% 
            add_p()      # P value for each strata
    ) 

```

## Specifying tables tests

```{r}
df_cint %>% 
    tbl_summary(by = sex,
                statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) %>% 
    add_p(
        test = list(
            all_continuous()~ "t.test",     # Specify T test for all continuous variables
            all_categorical() ~ "fisher.test"     # Specify fisher's text for all categorical variables
        )
    ) %>% 
    separate_p_footnotes()     # Specific p-value labelled

```

## Customised tables

```{r comment=NA}
df_cint %>% 
    select(bmi, hpt, bulb_0, sbp, bmicat) %>% 
    tbl_summary(
        by = hpt,
        statistic = list(
            bmi ~ "{mean} ({sd})",
            all_categorical() ~ "{n} ({p})",
            sbp ~ "{median} ({p25}, {p75})",
            bulb_0 ~ "{mean} ({sd})"),
         missing_text = "(Missing",
        digits = list(all_categorical() ~ c(0, 1))
    ) %>% 
    add_stat_label(label = list(
        bmi = "Mean(SD)",
        bulb_0 = "Mean(SD)",
        all_categorical() ~ "n(%)",
        sbp = "Median(IQR)"
    )) %>% 
    add_p(
        test = list(
            bmi ~ "t.test",
            bulb_0 ~ "t.test",
            all_categorical() ~ "fisher.test",
            sbp ~ "wilcox.test"
        )
    ) %>% 
    separate_p_footnotes()
```

## Creating data for the paired table

```{r}
df_paired_cint <- 
    df_cint_all %>% 
    select(sid, cart, cca_0, cca_12, ica_0, ica_12) %>%
    mutate(
        ica_12 = case_when(ica_12 > median(ica_12, na.rm=T) ~ "High",
                               ica_12 <= median(ica_12, na.rm=T) ~ "Low") %>% 
            factor(),
        ica_0 = case_when(ica_0 > median(ica_0, na.rm=T) ~ "High",
                               ica_0 <= median(ica_0, na.rm=T) ~ "Low") %>% 
            factor()
    ) %>% 
    select(sid, cart, cca_0, cca_12, ica_0, ica_12)

df_A <-
    df_paired_cint %>% 
    pivot_longer(cols =  c(cca_0, cca_12), 
                 names_to = c("cca", "period"), 
                 names_sep = "_", 
                 values_to  = "cca_measure") %>% 
    select(sid, cart, period, cca_measure)

df_paired_long <-
    df_paired_cint %>%
    pivot_longer(cols =  c(ica_0, ica_12), 
                 names_to = c("ica", "period"), 
                 names_sep = "_", 
                 values_to  = "ica_measure") %>% 
    select(sid, cart, period, ica_measure) %>% 
    full_join(df_A, by = c("sid", "period", "cart"))
```

## Paired table

```{r}
df_paired_long %>% 
    mutate(period = case_when(period == "0" ~ "Month 0",
                              period == "12" ~ "Month 12")) %>% 
    filter(complete.cases(.)) %>% 
    group_by(sid) %>% 
    filter(n()==2) %>% 
    ungroup() %>%
    tbl_strata(strata = cart, ~.x %>%
        tbl_summary(by = period, 
                    include = -sid,
                    statistic = list(cca_measure ~ "{mean} ({sd})",
                                     ica_measure ~ "{n} ({p})"),
                    label = list(ica_measure = "ICA",
                                 cca_measure = "CCA(mm)"),
                    digits = list(all_categorical() ~ c(0, 1)))%>% 
            add_p(test = list(ica_measure ~ "mcnemar.test",
                              cca_measure ~ "paired.t.test"),
                              group = sid,
                  pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
            separate_p_footnotes() %>% 
            add_stat_label(label = list(ica_measure = "n(%)",
                                        cca_measure = "mean(SD)"))
    ) %>% 
    modify_caption("**Table 2: Comparative month 0 and 12 measures**")
```

# Cross table

```{r}
df_cint %>% 
    gtsummary::tbl_cross(row = bmicat, 
                         col = sex, 
                         percent = "row", 
                         digits = c(0,1),
                         missing = "ifany",
                         missing_text = "[Missing]") %>% 
    add_p() %>% 
    modify_caption(caption = "**Table X: tbl_cross Example**")
```

# Regression models

Univariate regression

### Univariate linear regression

```{r comment=NA}
tbl1 <- 
    df_cint %>% 
    tbl_uvregression(y = sbp, method = lm) %>% 
    modify_header(update = list(estimate ~ "**Estimate**",
                                label ~ "**Variable**")) %>% 
    modify_caption(caption = "**Table XI:** Univariate linear regression")

tbl1 %>% show_header_names() # Show header names of table

tbl1
```

### Univariate logistic regression

```{r comment=NA}
df_mbu_clean %>% 
    select(-dura_adm) %>% 
    tbl_uvregression(
        method = glm,
        y = died,
        method.args = family(binomial),
        exponentiate = TRUE,
        pvalue_fun = function(x) style_pvalue(x, digits = 3)
    ) %>% 
    modify_header(update = list(estimate ~ "**Estimate**",
                                label ~ "**Variable**"))
```

### Univariate survival analysis

```{r comment=NA}
library(survival)

df_mbu_clean %>% 
    tbl_uvregression(
        method = coxph,
        y = Surv(dura_adm, died=="Yes"),
        exponentiate = TRUE,
        pvalue_fun = function(x) style_pvalue(x, digits = 3)
    ) %>% 
    modify_caption(caption = "**Table 10**: Univariate Cox's Proportional Model")
```

## Multivariate regression

### Multiple liner regreession

```{r comment=NA}
df_cint %>% 
    lm(sbp ~ ., data = .) %>% 
    tbl_regression(pvalue_fun = function(x) style_pvalue(x, digits = 3)) %>% 
    modify_header(update = list(estimate ~ "**Estimate**",
                                label ~ "**Variable**")) %>% 
    modify_caption(caption = "**Table XI:** Multivariate linear regression") %>%
    bold_labels() %>% 
    as_gt() %>% 
    gt::tab_options(
        table.font.size = 14,
        table.font.names = "Times New Roman",
        data_row.padding = 1
    ) 
```

### Multiple variable logistic regression

```{r}
df_mbu_clean %>% 
    glm(died ~ ., data = ., family = binomial) %>% 
    tbl_regression(exponentiate = T, include = -dura_adm) %>% 
    add_n(location = "level") %>% 
    add_nevent(location = "level") %>% 
    add_global_p() %>% 
    add_q() %>% 
    add_significance_stars(
        hide_p = FALSE,
        hide_ci = FALSE,
        hide_se = TRUE
    ) %>% 
    add_vif() %>% 
    modify_header(label = "**Predictor**") %>% 
    modify_caption(caption = "**Table 5: Highly customised logistic regression**") %>% 
    modify_footnote(ci = "CI = My 95%CI", abbreviation = TRUE) %>% 
    sort_p() %>% 
    bold_p(t=0.1, q=TRUE) %>% 
    bold_labels() %>% 
    italicize_levels()
    

```

### Cox proportional hazard regression

```{r}
df_mbu_clean %>% 
    coxph(Surv(dura_adm, died == "Yes")~., data = .) %>% 
    tbl_regression(exponentiate = T)
```

# Combining tables

## Merging

## Stacking

## Multiple comparison table

```{r}
library(titanic)
library(plotrix) #has a std.error function

# create smaller version of the dataset
df <- 
  titanic_train %>%
  select(Sex, Embarked, Age, Fare) %>%
  filter(Embarked != "") # deleting empty Embarked status

# first, write a little function to get the 2-way ANOVA p-values in a table
# function to get 2-way ANOVA p-values in tibble
twoway_p <- function(variable) {
  paste(variable, "~ Sex * Embarked") %>%
    as.formula() %>%
    aov(data = df) %>% 
    broom::tidy() %>%
    select(term, p.value) %>%
    filter(complete.cases(.)) %>%
    pivot_wider(names_from = term, values_from = p.value) %>%
    mutate(
      variable = .env$variable,
      row_type = "label"
    )
}

# add all results to a single table (will be merged with gtsummary table in next step)
twoway_results <-
  bind_rows(
    twoway_p("Age"),
    twoway_p("Fare")
  )
twoway_results



tbl <-
  # first build a stratified `tbl_summary()` table to get summary stats by two variables
  df %>%
  tbl_strata(
    strata =  Sex,
    .tbl_fun =
      ~.x %>%
      tbl_summary(
        by = Embarked,
        missing = "no",
        statistic = all_continuous() ~ "{mean} ({std.error})",
        digits = everything() ~ 1
      ) %>%
      modify_header(all_stat_cols() ~ "**{level}**")
  ) %>%
  # merge the 2way ANOVA results into tbl_summary table
  modify_table_body(
    ~.x %>%
      left_join(
        twoway_results,
        by = c("variable", "row_type")
      )
  ) %>%
  # by default the new columns are hidden, add a header to unhide them
  modify_header(list(
    Sex ~ "**Sex**", 
    Embarked ~ "**Embarked**", 
    `Sex:Embarked` ~ "**Sex * Embarked**"
  )) %>%
  # adding spanning header to analysis results
  modify_spanning_header(c(Sex, Embarked, `Sex:Embarked`) ~ "**Two-way ANOVA p-values**") %>%
  # format the p-values with a pvalue formatting function
  modify_fmt_fun(c(Sex, Embarked, `Sex:Embarked`) ~ style_pvalue) %>%
  # update the footnote to be nicer looking
  modify_footnote(all_stat_cols() ~ "Mean (SE)")
```

# hh

```{r}
# standardized difference
tbl1 <- 
  trial %>% 
  select(trt, age, marker) %>%
  tbl_summary(by = trt, missing = "no",
              statistic = all_continuous() ~ "{mean} ({sd})") %>%
  add_difference(all_continuous() ~ "cohens_d")

# table with p-value and corrected p-values
tbl2 <- 
  trial %>% 
  select(trt, age, marker) %>%
  tbl_summary(by = trt, missing = "no") %>%
  add_p(all_continuous() ~ "t.test") %>%
  add_q(method = "bonferroni") %>%
  modify_column_hide(all_stat_cols())

# merge tbls together
tbl_final <- 
  tbl_merge(list(tbl1, tbl2)) %>%
  # remove spanning headers
  modify_spanning_header(everything() ~ NA)

tbl1
tbl2
tbl_final
```

## References
