---
title: "Sample size"
---

```{r}
#| include: false
library(tidyverse)
```

In a study by Mlay et al, 2020, the mean weight gain and standard deviation for babies whose first
feed was within 48 hours of life was 14.3g/kg/day and 4.8. Mean weight gain and standard deviation
for babies whose first feed was more than 48 hours of life was 10.2g/kg/day, and 4.8. Using a Type 1
error ($\alpha$) of 0.05, a Type 2 error ($\beta$) of 20%, an average weight gain of 15g/kg/day in
those without the exposure, and 12.5g/kg/day for those with the exposure, and assuming the number of
neonates without the exposure will be twice those with exposure, we determine the sample size using
the formula by Chow et al below: $$
n_1 = r*n_2 \;\text{ and }\; n_2=\left(1+\frac{1}{r}\right) \left(\sigma\frac{z_{1-\alpha/2}+z_{1-\beta}}{\mu_1-\mu_2}\right)^2
$$ Where, $\mu_1$ and $\sigma_1$ are the mean and standard deviation of weight gain in infants
without exposure. $\mu_2$ and $\sigma_2$ are the mean and standard deviation of weight gain in
infants with exposure $\alpha$ is Type I error $\beta$ is Type II error, thus power is $1-\beta$ and
$r$ is the ratio of the numbers unexposed to the exposed.

Substituting into the formula:

```{r}
#| echo: false
mu1 = 12.5
mu2 =15
r = 2
sd = 4.8
alpha = 0.05
beta  = 0.20

n2 <- (1+1/r)*(sd*(qnorm(1-alpha/2)+qnorm(1-beta))/(mu1-mu2))^2

n1 <- r*n2

glue::glue(
    "The number of unexposed and exposed neonates required are ", 
    {ceiling(n1)}, " and ", {ceiling(n2)}, " respectively.")
glue::glue("Total sample size = ", {ceiling(n2)+ceiling(n1)})
```

```{r}
#| echo: false
n = ceiling(n2)+ceiling(n1)
z = 20
glue::glue("Adjusting for ", {z}, "% non-response rate we have a final sample size of ", {(n/(1-(z/100))) %>% ceiling()})
```
