---
title: "Survival Analysis"
---

```{r}
#| include: false

library(tidyverse)
```

## Reading and cleaning data

```{r comment=NA}
df_mbu <-
    readxl::read_xlsx("C:/Dataset/mbu.xlsx") %>% 
    janitor::clean_names() %>% 
    select(
        -c(pt_name, ip, diag1, diag2, diag3, address, 
              disweight, field1, del_date_time)) %>% 
    filter(
        (outcome %in% c("Died", "Discharged")) & 
        (wt < 8 & wt >= 1) & 
        (age < 29) & 
        (sex != "Missing") & 
        (place != "Missing") &
        (apgar1 < 11) &
        (apgar2 < 11) & 
        (!is.na(adm_date_time)) & 
        (!is.na(dis_date_time)) &
        (del != "Missing")) %>% 
    rename(apgar5 = apgar2,
           del_mode = del) %>% 
    mutate(gestage = ifelse(gestage > 50 | gestage < 26, NA, gestage),
           place = ifelse(place == "Self", "Home", place),
           place = ifelse(
               place == "Maternity Home", "Clinic/Hospital", place),
           across(c(sex, place), ~as.factor(.)),
           place = fct_relevel(
               place, c("KATH", "Clinic/Hospital", "Home")),
           died = outcome == "Died",
           adm_dura_hrs = difftime(
               dis_date_time, adm_date_time, units = "hours") %>% 
               as.numeric(),
           adm_dura_hrs = ifelse(
               adm_dura_hrs < 0 | adm_dura_hrs > 1040, NA, adm_dura_hrs),
           adm_year = format(adm_date_time, "%Y"),
           outcome = factor(outcome, levels = c("Discharged","Died")),
           del_mode = ifelse(
               del_mode %in% c("Forceps", "Vacuum"), 
               "Assisted VD", del_mode) %>% 
               factor(levels = c("SVD", "C/S", "Assisted VD"))) %>% 
    select(-c(adm_date_time, dis_date_time)) %>% 
    filter(adm_year != "2012" & !is.na(adm_dura_hrs))
```

## Labelling data

```{r comment=NA}
labelled::var_label(df_mbu) <-
    list(
        wt = "Weight in kgs",
        age = "Age in days",
        sex = "Sex",
        place = "Place of delivery",
        apgar1 = "First minute APGAR",
        apgar1 = "First minute APGAR",
        apgar5 = "Fifth minute APGAR",
        outcome = "Outcome",
        del_mode = "Mode of delivery",
        gestage = "Gestational age",
        died = "Baby died",
        adm_dura_hrs = "Duration of admission",
        adm_year = "Year of admission"
    )
```

## Summarizing data

```{r comment=NA}
df_mbu %>% 
    summarytools::dfSummary(graph.col = F, labels.col = FALSE)
```

## Summarizing by outcome

```{r comment=NA}
# gtsummary::theme_gtsummary_compact()
df_mbu %>% 
    gtsummary::tbl_summary(
        by = outcome,
        percent = "row",
        digits = list(gtsummary::all_categorical()~ c(0,1),
                      gtsummary::all_continuous()~0,
                      wt ~ 1)
    ) %>% 
    gtsummary::modify_spanning_header(
        gtsummary::all_stat_cols() ~ "Outcome"
        ) %>% 
    gtsummary::add_overall() %>%   
    gtsummary::add_p(
        pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
    gtsummary::bold_p() %>% 
    gtsummary::bold_labels()
```

## Overall Kaplan-Meier survival

```{r comment=NA}
km1 <- 
    df_mbu %>% 
    survival::survfit(
        survival::Surv(event = died, time = adm_dura_hrs) ~ 1, 
        data = .)
km1
```

```{r comment=NA}
km1 %>% broom::tidy()
```

```{r comment=NA}
km1 %>% 
    survminer::ggsurvplot(
        data = df_mbu,
        pval = T,
        risk.table = T, 
        censor = F,
        font.x = c(10, "bold.italic", "red"),
        font.y = c(10, "bold.italic", "darkred"),
        font.tickslab = c(10, "plain", "darkgreen"),
        tables.theme = survminer::theme_cleantable())
```

## Kaplan-Meier survival by sex

```{r comment=NA}
km_by_sex <- 
    df_mbu %>% 
    survival::survfit(
        survival::Surv(
            event = died, time = adm_dura_hrs) ~ sex, data = .)

km_by_sex
```

```{r comment=NA}
km_by_sex %>% broom::tidy() %>% arrange(time)
```

## Weekly survival

```{r}
summary(km_by_sex, times = seq(0, 1040, by = 24*7))
```

```{r comment=NA}
km_by_sex %>% 
    survminer::ggsurvplot(
        data = df_mbu,
        pval = T,
        risk.table = T, 
        censor = F,
        font.x = c(10, "bold.italic", "red"),
        font.y = c(10, "bold.italic", "darkred"),
        font.tickslab = c(10, "plain", "darkgreen"),
        tables.theme = survminer::theme_cleantable(),
        conf.int=T,
        tables.height = 0.2,
        break.x.by = 24*7,
        legend.title = "",
        surv.scale = "percent")
```

## Cox's regression

```{r comment=NA}
cox_by_sex <- 
    df_mbu %>% 
    survival::coxph(
        survival::Surv(
            event = died, time = adm_dura_hrs) ~ sex, data = .)

summary(cox_by_sex)
```

### Univariate Cox regression

```{r comment=NA}
cox_by_sex %>% 
    gtsummary::tbl_regression(
    exponentiate = TRUE,
    pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)
) %>% 
    gtsummary::bold_labels() %>% 
    gtsummary::bold_p() %>% 
    gtsummary::modify_caption("Univariate Cox regression with sex")
```

### Univariate Cox regression multiple variables at a time

```{r}
df_mbu %>% 
    gtsummary::tbl_uvregression(
        include = -outcome,
        method = survival::coxph, 
        y = survival::Surv(event = died, time = adm_dura_hrs),
        exponentiate = TRUE, 
        pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
    gtsummary::bold_p() %>% 
    gtsummary::bold_labels()
```

### Multivariate Cox regression

```{r comment=NA}
cox_1 <- 
    df_mbu %>% 
    select(-c(outcome, gestage)) %>% 
    survival::coxph(
        survival::Surv(event = died, time = adm_dura_hrs) ~ ., 
        data = .)

cox_2 <- 
    df_mbu %>% 
    select(-c(outcome)) %>% 
    survival::coxph(
        survival::Surv(event = died, time = adm_dura_hrs) ~ ., 
        data = .)
```

## Checking the model

```{r comment=NA}
performance::performance(cox_1)
performance::performance(cox_2)
```

## Compare two multivariate models: Obviously second is better than first

```{r comment=NA}
performance::compare_performance(cox_1, cox_2)
```

## Checking for proportional hazard assumption

Significant p-values indicates Proportional Hazard assumption violated

```{r comment=NA}
cox_1 %>% survival::cox.zph()
```

Residuals falling outside the standard error margins indicates Proportional Hazard assumption
violated. These violations will have to be dealt with.

```{r comment=NA}
cox_1 %>% survival::cox.zph() %>% survminer::ggcoxzph()
```

## Combining Univariate and multivariate Cox regression

### Without gestational age

```{r comment=NA}
cox_1A <-
    df_mbu %>% 
    select(-c(outcome, gestage)) %>%
    gtsummary::tbl_uvregression(
        method = survival::coxph, 
        y = survival::Surv(event = died, time = adm_dura_hrs),
        exponentiate = TRUE, 
        pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
    gtsummary::bold_labels() %>% 
    gtsummary::bold_p()

cox_1B <- 
    df_mbu %>% 
    select(-c(outcome, gestage)) %>% 
    survival::coxph(
        survival::Surv(
            event = died, time = adm_dura_hrs) ~ ., data = .) %>% 
    gtsummary::tbl_regression(
        exponentiate = TRUE,
        pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
    gtsummary::bold_labels() %>% 
    gtsummary::bold_p()

gtsummary::tbl_merge(
    tbls = list(cox_1A, cox_1B),
    tab_spanner = c("Univariate", "Multivariate")) %>% 
    gtsummary::modify_caption(
        "Combined Univariate and Multivariate Cox regression with sex")
```

### With gestational age

```{r comment=NA}
cox_2A <-
    df_mbu %>% 
    select(-c(outcome)) %>%
    gtsummary::tbl_uvregression(
        method = survival::coxph, 
        y = survival::Surv(event = died, time = adm_dura_hrs),
        exponentiate = TRUE, 
        pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3)) %>% 
    gtsummary::bold_labels() %>% 
    gtsummary::bold_p()

cox_2B <- 
    df_mbu %>% 
    select(-c(outcome)) %>% 
    survival::coxph(
        survival::Surv(
            event = died, time = adm_dura_hrs) ~ ., data = .) %>% 
    gtsummary::tbl_regression(
        pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = 3),
        exponentiate = TRUE) %>% 
    gtsummary::bold_labels() %>% 
    gtsummary::bold_p()

gtsummary::tbl_merge(
    tbls = list(cox_2A, cox_2B),
    tab_spanner = c("Univariate", "Multivariate")) %>% 
    gtsummary::modify_caption(
        "Combined Univariate and Multivariate Cox regression with sex")
```

## Plotting coefficients

```{r}
df_mbu %>% 
    select(-c(outcome)) %>% 
    survival::coxph(
        survival::Surv(event = died, time = adm_dura_hrs) ~ ., 
        data = .) %>% 
    sjPlot::plot_model(
        value.offset = 0.4,show.values = T, show.p = T) +
    geom_hline(yintercept = 1, alpha =.5, color = 'grey45') +
    labs(
        title = "Coefficients Plot of Hazard Ratios", 
        y = "Hazard ratio (95%CI)") +
    theme_minimal() +
    scale_y_log10(limits = c(0.4, 5))
```
